### 2017-01-11 robust-software.tex

Note: one practice \emph{not} included in the list below is the use of version control.
We think this is essential to building software of any kind
\cite{wilson2014,wilson2016}.
However,
our focus in this short guide is on the software that is built,
not on the process used to build it.
Similarly,
we do not explicitly discuss code review,
unit testing,
and other practices essential to maintainability and sustainable software development.

@@

\section{Use a build utility and package manager.}

In all but the simplest of scripts, \textbf{use a build
automation utility to compile code and/or manage your software workflow}. The benefits of
using tools like Make, Rake, Maven, Ant or MS Build vastly outweigh the
their learning curves. Most importantly, a build system
lets someone else build your software the same way you would. Most build
utilities can also be configured to run tests during every build and
re-generate documentation, and can be hooked into continuous integration,
cutting releases, and deploying the software to other environments.

Build utilities rely on any dependencies being present in the environment for a
successful build, and so developers should
\textbf{document \emph{all} dependencies, preferably in a machine-readable form}.
Package managers like apt and yum are available on most Unix-like systems, and
application package managers exist for specific languages like Python (pip),
Java (Maven/Gradle), and Ruby (RubyGems). These package managers can be used
together with the build utility to ensure that dependencies are available at
compile/run time.
For example, it is common for Python projects to include a file called
\texttt{requirements.txt} that lists the names of required libraries,
along with version ranges:

@@

This file can be read by the pip package manager, which can check that the
required software is available and install it if it is not. 
Whatever is used,
we recommend that developers should \emph{always} install dependencies
using their dependency description, especially on their personal machines, so that
they're sure it works.

>>>>

@@

\section{Rely on build tools and package managers for installation.}

Programmers routinely use build tools like Make, Rake, Maven, Ant or MS Build
to compile code, deploy applications, and automate other tasks.
These tools can also be used to manage runtime environments,
i.e.,
to check that the right versions of required packages are installed
and install or upgrade them if they are not.

The same tools can and should be used to manage runtime environments on users' machines as well.
Accordingly,
developed should
\textbf{document all dependencies in a machine-readable form}.
Package managers like apt and yum are available on most Unix-like systems, and
application package managers exist for specific languages like Python (pip),
Java (Maven/Gradle), and Ruby (RubyGems). These package managers can be used
together with the build utility to ensure that dependencies are available at
compile/run time.

For example, it is common for Python projects to include a file called
\texttt{requirements.txt} that lists the names of required libraries,
along with version ranges:

@@

This file can be read by the pip package manager, which can check that the
required software is available and install it if it is not. 
Whatever is used,
developers should \emph{always} install dependencies
using their dependency description, especially on their personal machines, so that
they're sure it works.
